---
title: "Data Analysis-Assignment 2"
  author: "Nawal Zehra Hasan"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Assignment 2 - Nawal Zehra Hasan 
##Introduction
For this assignment we were given the Hotels Europea Dataset [**Hotel Price dataset**](https://osf.io/p6tyr/download) with different features about the hotels that can be found here [**Hotel Features dataset**](https://osf.io/utwjs/download). By choosing one particular city, I wanted to understand how hotel ratings are associated with different features of the hotel such as distance and stars as the explanatory variables. 

##Understanding the filtered data
The table summary reveals that of the total of 667 observations 414 are highly rate(>4). The descriptive statistic tell us that of of the total of 667 observations 62% of the hotels are highly-rated. Besides, they are 2.75 miles away from city center on average, and they have 3.48 stars on average. These highly rated hotels are on average 1.58 miles away from the city center with  3.52 stars on average. 


```{r include=FALSE, message=FALSE, warning=FALSE}
# Clean Environment
rm(list=ls())
# Import libraries
library(tidyverse)
library(haven)
library(data.table)
library(rms)
library(lspline)
library(huxtable)
library(modelsummary)
install.packages("pscl")
library(pscl)
library(esquisse)
library(ggplot2)
library(fixest)
library(mfx)
```

```{r include=FALSE, message=FALSE, warning=FALSE}
# Load hotels Europe data sets
hotels_price <- read_csv ( "https://osf.io/p6tyr/download" )
hotels_features <- read_csv ( "https://osf.io/utwjs/download" )

# Join both by hotel_id
data <- left_join(hotels_price, hotels_features, by = "hotel_id" ) 
rm ( hotels_price , hotels_features )

#filter by year and city
unique ( data$city_actual )
unique ( data$year )


# Get the needed sample: hotels in Amsterdam
Amsterdam <- data %>% 
                filter (accommodation_type == "Hotel" ) %>%
                filter ( year == 2018 ) %>% 
                filter ( city_actual=="Amsterdam" ) %>% 
                filter ( !is.na ( rating ) , !is.na ( stars ), !is.na ( distance ) )


#creating binary variable for ratings
Amsterdam <- Amsterdam %>% 
                mutate ( highly_rated = ifelse(Amsterdam$rating >= 4, 1, 0 ) )

Amsterdam$log_price <- log ( Amsterdam$price )
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# Summary statistics on price and log of price
P95 <- function ( x ) { quantile ( x,.95,na.rm=T ) }
datasummary( highly_rated + distance + stars ~ Mean + SD + Min + Max + Median + P95 + N , data = Amsterdam, title = "Summary Statistics" )

#checking highly rated number of hotels
table ( Amsterdam$highly_rated )
```
```{r,  echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
#checking lowess for each independent variable
plot_dist <- ggplot(Amsterdam, aes(x=distance, y=highly_rated))+
  geom_smooth(formula = y~x, method = "loess")

plot_stars <- ggplot(Amsterdam, aes(x = stars, y = highly_rated ) ) +
  geom_smooth ( formula = y ~ x, method = "loess")

plot_price <- ggplot (Amsterdam, aes ( x = log ( price ), y = highly_rated ) ) +
  geom_smooth ( formula = y ~ x, method = "loess")
```

```{r, echo=FALSE, include=FALSE}
# Model 1:  LPM
lpm <- feols(highly_rated ~ lspline(distance,c(0.75, 3.5))  + stars + log_price , data=Amsterdam, vcov = "hetero")
Amsterdam$pred_lpm <- predict(lpm)
summary(Amsterdam$pred_lpm)

#I used the logit/probit models as the summarry statistics for LPM reveal that probability is > 1. This is unfit for predictions.

# Model 2: Logit
#logit coefficients
logit <- glm(highly_rated ~ lspline(distance,c(0.75, 3.5))  + stars + log_price , data=Amsterdam, family='binomial'(link = "logit"))
# Predicted probabilities 
Amsterdam$pred_logit <- predict.glm(logit, type="response")
# Marginal Differences
logit_marg <- logitmfx( highly_rated ~ lspline(distance,c(0.75, 3.5))  + stars + log_price , data=Amsterdam, atmean=F, robust = T)


# Model 3: Probit
#probit coefficients
probit <- glm(highly_rated ~  lspline(distance,c(0.75, 3.5))  + stars + log_price , data=Amsterdam, family=binomial(link="probit"))
# predicted probabilities 
Amsterdam$pred_probit<- predict.glm(probit, type="response") 
summary(Amsterdam$pred_probit)
# Marginal Differences
probit_marg <- probitmfx(highly_rated ~  lspline(distance,c(0.75, 3.5))  + stars + log_price , data=Amsterdam, atmean=F)
```

```{r}

pmodels <- list(lpm, logit, logit_marg, probit, probit_marg)
msummary( pmodels ,
          fmt="%.3f",
          gof_omit = 'DF|Deviance|Log.Lik.|F|R2 Adj.|AIC|BIC|R2|PseudoR2|Std.Errors',
          stars=c('*' = .05, '**' = .01),
          title = "Probability of Highly Rated Hotels by Distance and Stars- LPM, Logit & probit MODELS"
          )
```

