---
title: "Data Analysis 2 - Assignment 1"
author: "Nawal Zehra Hasan"
  pdf_document:
    extra_dependencies: ["float"]
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
#cleaning the environment
rm ( list = ls ( ) )
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(readr)
library(estimatr)
library(huxtable)
library(dplyr)
library(knitr)
library(ggplot2)
install.packages("modelsummary")
library(modelsummary)
```
### Data Analysis 2 - Assignment 1
## Nawal Zehra Hasan
## 27/11/2021

### Introduction
Wages, Sex, and Highest Grade Completed: I expect to see an impact on wage gap and the highest degree completed. I want to explore whether this difference prevails regardless of gender. The question here would be, does it matter whether you are male or female if the highest degree you have completed is the same across both genders? 

As an analyst I was assigned a task to perform a comparative analysis of differences in the hourly wage rate received by men and women as Chief Executives. I filtered the data as per my occupation to be able to conduct an accurate analysis. I included education level bachelors and above, age to be 35 and above and working hours to be 40 and above. I also created dummy variables for education degrees and calculated the natural log of wages using hourly wages. 

### Analysis & Interpretation
To see the unconditional gender gap in wages I used summary statistics first to give an overview. I also created a histogram that reveals an evident gap in wages in between the two genders in our sample. Next, I modeled this as a regression to view and analyse this in our regressions table. 

I used lm robust regression to model my variables and show interactions as robust provides a weighting mechanism and weighs down influential observations. I also used robust as there is no compelling reason to exclude any extreme value in my observations. It also helped give robust standard errors and conduct coefficients tests. I ran a total of 5 regressions, all of which are using log of hourly wages hence a log-level regression.

Interpretation & Analysis

In the sample taken with the occupation Chief Executives we have a total of 874 observations. Out of these 211 which makes around 24% are females.

To check the unconditional gender gap, I used hourly wages for men and women. I also used log of hourly wages to show the gap and compare it with our regression model(reg1). 

Regression 1 in figure 6 reveals that controlling for all other factors, females as Chief Executives tend to earn 10.34% less than the men working as Chief Executives. It has a SE of 0.04 giving a confidence interval of [-0.18, -0.03]. This p-value of this coefficient is less than 0.01 as indicated by the asterisks. This indicates that the coefficient is statistically significant at the 0.01 level. We can also check the statistical significance by looking at the confidence interval. As the confidence interval does not contain the null hypothesis value i.e. 0, the coefficient is statistically significant. The intercept 3.7837 shows the average log earnings for males i.e. baseline group, in the filtered dataset.

Regression 2 is modeled using all the other education degrees except Bachelor, making bachelor our baseline. With a PhD which is the highest degree in our sample, our observations will earn only 0.5% more than those with a bachelors. However, our coefficient is not statistically significant, and we fail to reject our null hypothesis. We can possibly conclude that increasing levels of education beyond the bachelor's degree are correlated with higher wage levels, but only up to a certain point. While we see the coefficient for the dummy variable on Masters is statistically significant, the case is not the same with the dummy variable on PhD. We can possibly conclude that increasing levels of education beyond the bachelor's degree are correlated with higher wage levels, but only up to a certain point. While we see the coefficient for the dummy variable on Masters is statistically significant, the case is not the same with the dummy variable on PhD.

One study indicates that the wage gap widens with a college degree, that is wage gap between men and women with a college degree is wider than the wage gap between men and women without a college degree, therefore I added an interaction term in the Regression 3 model for education and gender. The coefficient for female:ed_MA suggests that the wage gap between men and women of with an education level of Master’s is, on average, 10.29% more than the wage gap for our baseline education level i.e. Bachelor’s. The coefficient for female:ed_PhD suggests that the wage gap between men and women of with an education level of PhD is, on average, 20.64% more than the wage gap for our baseline education level i.e. Bachelor’s. This might be an indication that with higher education levels, the wage gap between men and women seems to increase. However, since our coefficients for the interaction terms are statistically insignificant, we cannot me confident in these estimates.

Regression 4 shows that on average, controlling for other factors, a one year increase in age corresponds to a %0.03 increase in log earnings. We can see that the coefficient for age is statistically significant at a p < 0.05. By taking age females earn around 10.4% less than their male counterparts. 

Summary
Although all our regression models depict that women overall earn less than male with education and age but this does not necessarily mean discrimination against women as there are other factors that can be taken into consideration.


### Figures
```{r, echo=FALSE, warning=FALSE, message=FALSE}
#Loading the data-cps-earnings data set
df <- read_csv ( "https://osf.io/4ay9x/download" )

#check the first few rows of the data set
head ( df )

#no. of women in the data set
no_females <- count ( df [ which ( df$sex == 2 ) , ] ) #73738/149316 are women approx 49%

#selecting relevant observations
data <- df %>% filter ( occ2012 == 10, #code for chief executives
                    grade92 >= 43, #edu. level Masters or above
                    uhours >= 40, #working hours
                    earnwke >= 0, #weekly earnings
                    age >= 35 ) #35 or above years of age

#no. of women in the occupation                                
no_females <- count ( data [ which ( data$sex == 2 ) , ] ) # 211/874 females approx 24%

#Creating relevant variables to the data set
data <- data %>% mutate ( female=as.numeric(sex == 2 ) ) %>%
                 mutate ( h_wage = earnwke/uhours ) %>%
                 mutate ( lnw = log ( h_wage ) ) %>% 
                 mutate ( ed_BA=as.numeric(grade92==43),
                          ed_MA=as.numeric(grade92==44),
                          ed_Profess = as.numeric(grade92==45),
                          ed_PhD = as.numeric(grade92==46 ) )
```

#Unconditional gender gap
```{r, echo=FALSE, warning=FALSE, message=FALSE}

data %>% dplyr::select(earnwke,uhours,h_wage) %>% summary()
tabulate(data$female)
table(data$occ2012,data$female)

#recode male-female
data <- data %>%  mutate ( sex = recode_factor ( sex,
                                `1` = "Male",
                                `2` = "Female" ) )
#summary statistics with hourly wages
datasummary(sex*h_wage ~ 
               Mean + SD + Min + Max + N , data = data )

#summary statistics with log wages
datasummary(sex*lnw ~ 
               Mean + SD + Min + Max + N , data = data )

```

#wage gap between male and female
```{r, echo=FALSE, include = FALSE, fig.width= 3, fig.height= 3 }
#unconditonal gender gap by hourly wages
ggplot(data, aes(lnw, fill = sex)) + geom_histogram(alpha = 0.5, position = 'identity') +
  labs (x = "Log Hourly Wage", y = "Number of Respondents")

```
 
```{r, echo=FALSE, message=FALSE, warning=FALSE} 
# log-level regressions
#uncondtional gender gap
reg1 <- lm_robust(lnw ~ female, data = data, se_type = "HC1")
summary(reg1)
#with all education levels except BA
reg2 <- lm_robust(lnw ~ female + ed_MA + ed_PhD + ed_Profess, data = data, se_type = "HC1")
summary(reg2)
#with interactions between education degree and gender
reg3 <- lm_robust(lnw ~ female + ed_MA + ed_Profess + ed_PhD + female*ed_MA + female*ed_Profess + female*ed_PhD, data = data, se_type = "HC1")
summary(reg3)
#adding age as an explanatory variable with education level
reg4 <- lm_robust(lnw ~ female + age + ed_BA + ed_MA + ed_Profess + ed_PhD, data = data, se_type = "HC1")
summary(reg4)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE} 
#regression table for analysis
huxreg(reg1, reg2, reg3, reg4, statistics = c(N = "nobs", R2 = "r.squared"), stars = c(`****` = 0.001, `***` = 0.01, `**` = 0.05, `*` = 0.1 ),borders = 0.4, outer_borders = 0.8, number_format = "%.4f", align = ".")

```
```{r, echo=FALSE,message=FALSE, warning=FALSE, fig.width=7, fig.height=3}
ggplot(data = data, aes(x = grade92, y = lnw , colour = sex )) + geom_point() + stat_smooth(method = 'loess', formula = y ~ x, size = 1) + theme(axis.text.x=element_text(angle = 0, vjust = 0.5)) + xlim(40,50) + 
  labs( x = "Highest Degree Completed", y = "Log of Hourly Wages") +
 scale_x_continuous(labels = c("ed_BA", "ed_MA", "ed_Profess", "ed_PhD") )
  
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
data <- data %>% mutate ( Degree = recode_factor(grade92,
                                             `43` = "Bachelors",
                                             `44` = "Masters",
                                             `45` = "Professional",
                                             `46` = "Doctorate")) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=3  }
#education level of chief executives
ggplot(data, aes(x=Degree, fill=Degree)) +
  geom_histogram(stat="count") +
  scale_fill_manual(values=c('tan1','darkgray', 'wheat4', '#DDAD4B'))
```


